Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-05-24 paperDistance/paper.tex
minor edits and comments

We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust. With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way. The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference. Dust corrections are particularly non trivial for stars with poor parallax signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion). Instead, we correct for dust using a point estimate from the 3D dust map (Schlafly+Finkbeiner) sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$. In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior, we infer more precise parallaxes to all the stars in our sample. We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star. And repeat. This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd\ and bias the parallax inference. For each distance, we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD). We then regenerate our prior using the lightly dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution. We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

@@

@@

Transition clause/sentence. The biggest limitation of our method for generating the \cmd\ from the data is that we made no use of any kind of \tgas\ selection
function or completeness estimate or inverse-selection-volume corrections.
This model is a model of the \tgas\ Catalog
(or really the \tgas--\tmass\ intersection), not of the stars in
the Milky Way, nor any volume-limited subsample of the Milky Way.
Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
intersection, the model \emph{is} representive of the full intersection,
even the distant parts, where all the parallaxes are measured at low
signal-to-noise.
This is in contrast to techniques that might build the model from only
the high signal-to-noise sources; such a model would be biased towards
stellar types and compositions found locally, as well as lower-luminosity stars,
and stars which happened to get good parallaxes.
However, our use of the entire \tgas--\tmass\ intersection without
any selection function restricts the use of our \cmd\ model as a prior
to inferences within that same \tgas--\tmass\ intersection.
It would be biasing to use this same prior to infer
distances for the full billion-star catalog released in \gaia\ DR1
alongside \tgas, or for stars fainter or brighter or bluer than those
included in the \tgas--\tmass\ intersection.

@@

@@

This project was partially supported by This research was partially supported by
  the \acronym{NSF} (\acronym{AST-1517237}),
  \acronym{NASA} (grant \acronym{NNX12AI50G}),
  and the Moore-Sloan Data Science Environment at \acronym{NYU}.
It made use of the \acronym{NASA} Astrophysics Data System.
All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license.

>>>>

We need to correct the photometry used in our method, both for the generation of the prior and the likelihood, for dust. With the emergence of 3D dust maps, it is now possible to estimate dust corrections to stars within the Milky Way. The challenge is getting a proper estimate of the distance to the star before inferring $\varpi_{\true}$, since we need a dust correction prior to doing our inference. Dust corrections are particularly non trivial for stars with poor parallax signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. Although the dust model is probabilistic, as well as our distance inference, to move forward, we break from a fully Bayesian framework, in which we might infer the distance and the dust simultaneously (more about this in the discussion). Instead, we correct for dust using a point estimate from the 3D dust map \citep{green15} sampled at a distance estimated from iteratively inferring and then sampling our posterior pdf over $\varpi_{\true}$. In more detail, to apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior, we infer more precise parallaxes to all the stars in our sample. We then use this more precise parallax posterior to get a measurement of dust in the 3D dust map for each star. And repeat. This approach minimizes severe distances, and their associated severe dust corrections, which would occur if using the likelihoods alone. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median, for example, to do a light dust correction, because over correcting for dust can shoot the likelihood into an unphysical, different space of the \cmd\ and bias the parallax inference. For each distance, we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD). We then regenerate our prior using the lightly dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads, again, to a lighter deconvolution. \BLedit{In other words, the features of our inferred \cmd\ are conservatively broad (broader than if we over-estimated the errors which would have lead to excessive deconvolution and unrealistically narrow features).} We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

@@

\BLcomment{Also add the statistics about the parallax uncertainty shrinkage. Or put in a separate subsection? In fact, I now realize some of the figures are not discussed in the text. }

@@

Transition clause/sentence. The biggest limitation of our method for generating the \cmd\ from the data is that we made no use of any kind of \tgas\ selection
function or completeness estimate or inverse-selection-volume corrections.
This model is a model of the \tgas\ Catalog
(or really the \tgas--\tmass\ intersection), not of the stars in
the Milky Way, nor any volume-limited subsample of the Milky Way.
Importantly, because we used \emph{all} of the stars in the \tgas--\tmass\
intersection, the model \emph{is} representive of the full intersection,
even the distant parts, where all the parallaxes are measured at low
signal-to-noise.
This is in contrast to techniques that might build the model from only
the high signal-to-noise sources; such a model would be biased towards
stellar types and compositions found locally, as well as lower-luminosity stars,
and stars which happened to get good parallaxes.
However, our use of the entire \tgas--\tmass\ intersection without
any selection function restricts the use of our \cmd\ model as a prior
to inferences within that same \tgas--\tmass\ intersection.
It would be biasing to use this same prior to infer
distances for the full billion-star catalog released in \gaia\ DR1
alongside \tgas, or for stars fainter or brighter or bluer than those
included in the \tgas--\tmass\ intersection.
\BLcomment{I would specifically mention again the assumption of stationarity, and explain how it is broken by the selection function (in particular fluctuations in the detection probability of objects as a function of position, parallax, photometry, etc). Here, or in the critical discussion of assumptions.}

@@

\BLcomment{it's weird to end with these subsections. Cross-validation should go in the methods part (I think there is a flag there already), and the discussion of assumptions should go in the discussion. There should be a conclusion and look towards the future.}

@@

This project was partially supported by This research was partially supported by
  the \acronym{NSF} (\acronym{AST-1517237}),
  \acronym{NASA} (grant \acronym{NNX12AI50G}),
  and the Moore-Sloan Data Science Environment at \acronym{NYU}.
It made use of the \acronym{NASA} Astrophysics Data System.
All the code used in this project is available online\footnote{\url{https://github.com/andersdot/stellarTwins}} under an open-source license. \BLcomment{wrong link.}
